#!/usr/bin/env KnitWrapper.sh

---
title: "Simulation of an equity"
author: "Yvan Lengwiler"
date: "2025-09-11"
license: "MIT"
output:
    html_document: default
    word_document: default
    pdf_document: default
---

This script downloads some stock market data from finance.yahoo.com and then performs a Monte Carlo simulation.

```{r setup, include=FALSE}
# install and load some packages if they are missing
packages <- c('yfR','scales','here','curl') 
missing_packages <- !(packages %in% rownames(installed.packages()))
if (any(missing_packages)) {install.packages(packages[missing_packages])}
invisible(lapply(packages, library, character.only = TRUE))

# select location of this file as working directory
setwd(here())
```

# 1 Parameters

We begin by setting some parameters. These parameters will determine what the script does exactly:

```{r, set parameters, include=FALSE}
#symbol    <- 'AMD'              # Advanced Micro Devices
#symbol    <- 'NVDA'             # NVIDIA Corp
#symbol    <- 'UBS'              # UBS Group, NYSE
symbol    <- 'UBSG.SW'          # UBS Group, SIX
#symbol    <- 'GOOG'             # Alphabet
#symbol    <- 'NESN.SW'          # Nestlé
#symbol    <- 'GM'               # General Motors
#symbol    <- 'MSFT'             # Microsoft
#symbol    <- 'PM'               # Philip Morris
#symbol    <- 'XOM'              # Exxon Mobil
#symbol    <- '^GSPC'            # S&P500 index
#symbol    <- 'BTC-USD'          # BTC in USD
#symbol    <- 'CHFUSD=X'         # CHF-USD FX

from_date  <- '2010-12-01'
to_date    <- '2024-12-31'
interval   <- 'monthly'

sim_years  <- 2                 # length of simulation in years
n          <- 1000              # number of simulations

draw_paths <- TRUE              # make charts of not?
```

```{r, helper functions, include=FALSE}
# we will use this later
factor <- switch(
    interval,
        'monthly' = 12,          # number of trading months
        'weekly'  = 365.25/7,    # number of trading weeks
        'daily'   = 365.25*5/7   # number of trading days
    )
```

```{r, results='asis', echo=FALSE}
cat(sprintf('Data from %s is used from %s to %s in a %s interval.\n', symbol, from_date, to_date, interval))
cat(sprintf('This is then used to generate %i random paths for %#.2f years into the future.\n', n, sim_years))
```

The data are downloaded now, but the code for that is hidden.

```{r, download data, include=FALSE}
data <- yf_get(
    tickers = symbol,
    first_date = from_date,
    last_date = to_date,
    freq_data = interval,
    do_cache = TRUE,
    be_quiet = TRUE
)
price <- as.numeric(data$price_adjusted)
dates <- as.Date(data$ref_date)
yield  <- diff(log(price)) * factor
```

# 2 Compute Monte-Carlo simulation assuming normal distribution

Write code that generates random future paths of the equity price. Your
simulation should cover 'sim_years' years and you should generate 'n' such
paths.

The following functions are helpful in this endeavor:
'rnorm' : This function generates normally distributed random numbers.
'apply' : This procedure applies a function to each element of a matrix.
'cumsum': Compute the cumulative sum of a vector, i.e. x1, x1+x2, 
          x1+x2+x3, ..., x1+x2+x3+...+xn.
You can use ?command (e.g. ?rnorm) to get some help for any command.

===> WORK FOR YOU HERE <===

```{r, compute MC}
# first and second moment of the data
mu <- mean(yield)
sigma <- sd(yield)

# how many steps do we have to simulate?
# ...

# compute random returns
# ...

# accumulate random returns to get log levels
# ...

# transform log level to plain level
# ...
```

# **** make some charts *******************************************************

# ===> WORK FOR YOU HERE <===

# ...

