#!/usr/bin/env KnitWrapper.sh

---
title: "Estimation of a GARCH process"
author: "Your Name"
date: "2025-09-07"
license: "MIT"
output: pdf_document
geometry: a4paper
fontsize: 12pt
fontfamily: charter
---

This script downloads some stock market data from finance.yahoo.com and
them fits a GARCH(1,1) to the data. It also produces some informative charts.

```{r setup, include=FALSE}
packages <- c('yfR','scales','here','curl','rstudioapi') 
missing_packages <- !(packages %in% rownames(installed.packages()))
if (any(missing_packages)) {install.packages(packages[missing_packages])}
invisible(lapply(packages, library, character.only = TRUE))

# select location of this file as working directory
setwd(here())
```

# 1 Parameters

We begin by setting some parameters. These parameters will determine what the script does exactly.

```{r, set parameters, echo=FALSE}
#symbol    <- 'AMD'              # Advanced Micro Devices
#symbol    <- 'AMD'              # Advanced Micro Devices
#symbol    <- 'UBS'              # UBS Group, NYSE
symbol    <- 'UBSG.SW'          # UBS Group, SIX
#symbol    <- 'GOOG'             # Alphabet
#symbol    <- 'NESN.SW'          # Nestlé
#symbol    <- 'GM'               # General Motors
#symbol    <- 'MSFT'             # Microsoft
#symbol    <- 'PM'               # Philip Morris
#symbol    <- 'XOM'              # Exxon Mobil
#symbol    <- '^GSPC'            # S&P500 index
#symbol    <- 'BTC-USD'          # BTC in USD
#symbol    <- 'CHFUSD=X'         # CHF-USD FX

from_date <- '2010-12-01'
to_date   <- '2023-12-31'
from_date <- '2005-01-01'
to_date   <- '2025-08-31'
interval  <- 'daily'

# number of observations per year
factor <- switch(
    interval,
        'monthly' = 12,          # number of trading months
        'weekly'  = 365.25/7,    # number of trading weeks
        'daily'   = 365.25*5/7   # number of trading days
    )
```

# 2 Acquire data

We now download the relevant data from `https://finance.yahoo.com`. We use the `yfR` library for this.

```{r, download data, echo=FALSE}
data <- yf_get(
    tickers = symbol,
    first_date = from_date,
    last_date = to_date,
    freq_data = interval,
    do_cache = TRUE,
    be_quiet = TRUE
)
price <- as.numeric(data$price_adjusted)
dates <- as.Date(data$ref_date)

yield <- diff(log(price)) * factor
```

# 3 Estimate GARCH

## 3.1 Define likelihood function

param is a vector of four parameters: omega, alpha, beta, and initial_sigma2
(this is the variance at the first date of the sample).

An initial param has to be supplied for the estimation. The estimation
algorithm returns an optimized parameter vector.

The 'GARCH' function computes the log likelihood of the distribution implied
by the parameters for each point in time. It returns this vector of log
likelihoods (`logL`), the residuals (`eps`), and the variances (`sigma2`) for
each point in time.

'negLL' uses GARCH and then sums `logL` and multiplies it with -1, because
the optimizer searches for a minimum, not a maximum.

Finally, 'nlminb' is the algorithm we use to find the best fit. This is a
non-linear optimizer based on Newton-Raphson that allows us to specify
bounds for the parameters. We want `0<omega`, `0<alpha<1`, and `0<beta<1`, and set
`initial_sigma2` equal to the unconditional variance `omega / (1 - alpha - beta)`.
 

```{r, likelihood function}
# log likelihood function
GARCH <- function(param, eps) {
  
    # give names to parameters
    omega       <- param[1]
    alpha       <- param[2]
    beta        <- param[3]
   
    # WORK FOR YOU HERE
    # Program the GARCH process for sigma2 here and compute the log likelihood
    # for each observed period.
    
    # ...
    
    # return results
    # These should be two vectors, with one entry for each observed date.
    return(list("logL"=logL, "sigma2"=sigma2))
}

# objective function: negative sum of period log densities
negLL <- function(param, eps) {
    result <- GARCH(param, eps) 
    return(-sum(result$logL))
}
```

## 3.2 Perform the estiation

We set the starting point for non-linear optimization first

Note: experience dictates that alpha is normally small and beta is often
rather large, so this is where we start.

```{r, starting point}
param <- c(
    var(yield) * (1 - 0.9),  # omega
    0.1,                     # alpha
    0.8,                     # beta
    var(yield)               # initial sigma2
)
```

```{r, running non-linear optimization}
    # WORK FOR YOU HERE
    # running the NL opt (use the nlminn procedure here)
    # ...
```

# 4 Reporting results

```{r, collect GARCH results, echo=FALSE}

if (out$convergence != 0) {
    cat('Estimation did not converge!')
    cat(out$message)
    stop()
}

# extract optimized parameters
omega       <- out$par[1]
alpha       <- out$par[2]
beta        <- out$par[3]
init_sigma2 <- out$par[4]

if (alpha+beta >= 1) {
    cat('Variance is not stable.')
    stop()
}

# extract eps and sigma of optimized process
result      <- GARCH(out$par, yield)
sigma2      <- result$sigma2
sigma       <- sqrt(sigma2)
```

Asset: `r symbol` (from `r from_date` to `r to_date`)

Estimation converged in `r out$iterations` iterations.

`r cat(out$message)`

neg likelihood (minimized) = `r out$objective`

mu = `r mu`

omega = `r omega`

alpha = `r alpha`

beta = `r beta`

1-alpha-beta = `r 1-alpha-beta`

initial volatility (t=1) = `r sqrt(init_sigma2)`

unconditional volatility = `r sqrt(omega/(1-alpha-beta))`

` `

# 5 Make some charts

    # WORK FOR YOU HERE
    # ...

